---
title: Election Fraud in North Carolina
author: ~
date: '2019-04-17'
slug: election-fraud-in-north-carolina
categories: ["R"]
tags: []
type: post
description: description
keywords:
  - key
  - words
topics:
  - topic 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 12, fig.height = 6)
library(tidyverse)
library(magrittr)
library(reshape2)
library(RColorBrewer)
library(knitr)
```

# Prep Work

## Load Election Data

This project will be working with the precinct by precinct results of the 11/06/2018 midterm election in the state of North Carolina. These results were obtained from the North Carolina State Board of Elections & Ethics Enforcement website on 11/30/2018 at the following URL: https://dl.ncsbe.gov/ENRS/2018_11_06/results_pct_20181106.zip
```{r data-in, message = FALSE, warning=FALSE}
nc_results <- read_delim(
  "C:/datasets/results_pct_20181106.txt", "\t",
  escape_double = FALSE, trim_ws = TRUE
)

names(nc_results) <- names(nc_results) %>% 
  tolower() %>%
  str_replace_all(" ", "_")

nc_results <- nc_results %>%
  mutate(
    choice = if_else(choice == "" | is.na(choice), "NONE", choice),
    choice_party = if_else(
      choice_party == "" | is.na(choice_party),
      "NONE",
      choice_party
    )
  ) %>%
  mutate_if(is.numeric, as.integer) %>%
  rename(early = one_stop, mail = absentee_by_mail)
```

## Create Tables

### District 9

```{r d9}
house_d9 <- nc_results %>%
  filter(contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09")

d9_candidate_by_county <- house_d9 %>%
  group_by(county, choice) %>%
  summarise_at(vars("total_votes", "election_day", "early", "mail", "provisional"), sum) %>%
  ungroup()

d9_county_totals <- d9_candidate_by_county %>%
  group_by(county) %>%
  select(-choice) %>%
  summarise_all(sum) %>%
  ungroup()

d9_by_county <-
  bind_rows(
    d9_candidate_by_county,
    d9_county_totals %>% mutate(choice = "county_total")
  ) %>%
  arrange(county)
```

### All North Carolina US House Races
```{r nc-house-races}
nc_house_all <- nc_results %>%
  filter(str_detect(contest_name, "US HOUSE OF REPRESENTATIVES")) %>%
  mutate_at(c("contest_name", "choice", "choice_party"), factor)

all_house_candiate_by_county <- nc_house_all %>%
  group_by(contest_name, county, choice, choice_party) %>%
  summarise_at(vars("total_votes", "election_day", "early", "mail", "provisional"), sum) %>%
  ungroup()

house_county_totals <- all_house_candiate_by_county %>%
  group_by(contest_name, county) %>%
  select(-choice, -choice_party) %>%
  summarise_all(sum) %>%
  ungroup()

house_by_county <-
  bind_rows(
    all_house_candiate_by_county %>%
      mutate_at(c("county", "contest_name", "choice", "choice_party"), as.character),
    house_county_totals %>%
      mutate(county = as.character(county), contest_name = as.character(contest_name), choice = "county_total", choice_party = "county_total")
  ) %>%
  arrange(contest_name, county) %>%
  mutate_at(c("county", "contest_name", "choice", "choice_party"), factor)
```

## Set up Colors

```{r colorsetup}
red_blue_green <- brewer.pal(3, "Set1")
red <- red_blue_green[1]
blue <- red_blue_green[2]
green <- red_blue_green[3]

d9_color_scale <-
  scale_fill_manual(
    name = "Candidate",
    values = setNames(red_blue_green, c("Mark Harris", "Dan McCready", "Jeff Scott"))
  )

party_color_scale <-
  scale_fill_manual(
    name = "Party",
    values = setNames(red_blue_green, c("REP", "DEM", "OTHER"))
  )
```

# The 9th Congressional District

## Results of the Race for House of Representatives

```{r district9overall,results="hold",collapse=TRUE}
table <- d9_by_county %>%
  filter(choice != "county_total") %>%
  group_by(choice) %>%
  summarise(votes = sum(total_votes))

ggplot(data = table) +
  geom_col(aes(x = fct_reorder(choice, votes), y = votes, fill = choice)) +
  d9_color_scale +
  xlab("Candidate") +
  ylab("Votes") +
  scale_y_continuous(labels = scales::comma_format()) +
  ggtitle("Overall Results")

table %>%
  arrange(votes) %>%
  mutate(percent = votes / sum(votes)) %>%
  kable(caption = "9th Congressional District Results")
```

## County Level Results

```{r district9counties, results="hold"}
county_order <- d9_county_totals %>%
  arrange(total_votes) %>%
  pull(county) %>%
  as.character()

table <- d9_candidate_by_county %>%
  mutate(county = factor(county, levels = county_order))

ggplot(data = table) +
  geom_col(aes(x = county, y = total_votes, fill = choice)) +
  d9_color_scale +
  scale_y_continuous(labels = scales::comma_format()) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("County")+
  ylab("Votes")+
  ggtitle("County Vote Totals")

ggplot(
  data = table,
  aes(x = county, weight = total_votes, fill = choice)
) +
  d9_color_scale +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Share of Votes", x = "County") +
  geom_hline(yintercept = .5) +
  ggtitle("County Vote Split")

table <-
  d9_county_totals %>%
  gather(
    key = vote_type,
    value = vote_count,
    election_day,
    early,
    mail,
    provisional
  ) %>%
  mutate(county = factor(county, levels = county_order))

ggplot(
  table %>% select(-total_votes),
  aes(x = county, y = vote_count, fill = vote_type),
  position_stack()
) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("County") +
  ylab("Votes") +
  ggtitle("Breakdown of Voting Methods by County")

ggplot(
  table %>% select(-total_votes),
  aes(x = county, weight = vote_count, fill = vote_type)
) +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Percentage of Votes Cast", x = "County") +
  ggtitle("Proportion of Voting Methods by County")
```

## 9th District Vote by Mail vs Voting in Person 

```{r d9-mailin-inperson}
table <- d9_by_county %>%
  mutate(in_person = total_votes - mail) %>%
  gather(
    key = vote_type,
    value = vote_count,
    total_votes,
    election_day,
    early,
    mail,
    provisional,
    in_person
  ) %>%
  filter(!choice == "county_total") %>%
  filter(vote_type == "mail" | vote_type == "in_person") %>%
  group_by(county, vote_type) %>%
  mutate(county_total = sum(vote_count)) %>%
  mutate(share = vote_count / county_total) %>%
  select(-vote_count, -county_total) %>%
  filter(choice == "Dan McCready") %>%
  spread(vote_type, share) %>%
  ungroup() %>%
  mutate(county = factor(county, levels = county_order))

ggplot(data = table) +
  geom_col(
    aes(x = county, y = mail - in_person, fill = in_person < mail)
  ) +
  scale_fill_manual(values = setNames(c(blue, red), c(TRUE, FALSE))) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(y = "Share of Vote by Mail vs Share of In Person Votes") +
  ggtitle("Partisan Lean of Vote by Mail vs in Person Voting")
```

# State Wide Trends

## Overview of US House of Representatives Races

```{r houseTable,results="hold",collapse=TRUE}
statewide_shares <- all_house_candiate_by_county %>%
  mutate(choice_party = if_else(choice_party == "DEM" | choice_party == "REP", as.character(choice_party), "OTHER")) %>%
  group_by(choice_party) %>%
  summarise(party_early = sum(early), party_election_day = sum(election_day), party_mail = sum(mail), party_in_person = sum(total_votes - mail), party_total = sum(total_votes)) %>%
  ungroup() %>%
  mutate(
    party_election_day_share = party_election_day / sum(party_election_day),
    party_early_share = party_early / sum(party_early),
    party_mail_share = party_mail / sum(party_mail),
    party_in_person_share = party_in_person / sum(party_in_person),
    party_total_share = party_total / sum(party_total)
  ) %>%
  mutate(choice_party = fct_reorder(choice_party, party_total)) %>%
  arrange(party_total)

statewide_shares_no_uncontested <- all_house_candiate_by_county %>%
  filter(contest_name != "US HOUSE OF REPRESENTATIVES DISTRICT 03") %>%
  mutate(choice_party = if_else(choice_party == "DEM" | choice_party == "REP", as.character(choice_party), "OTHER")) %>%
  group_by(choice_party) %>%
  summarise(party_early = sum(early), party_election_day = sum(election_day), party_mail = sum(mail), party_in_person = sum(total_votes - mail), party_total = sum(total_votes)) %>%
  ungroup() %>%
  mutate(party_election_day_share = party_election_day / sum(party_election_day), party_early_share = party_early / sum(party_early), party_mail_share = party_mail / sum(party_mail), party_in_person_share = party_in_person / sum(party_in_person), party_total_share = party_total / sum(party_total)) %>%
  mutate(choice_party = fct_reorder(choice_party, party_total)) %>%
  arrange(party_total)

statewide_shares %>% kable(caption = "All House Races")
statewide_shares_no_uncontested %>% kable(caption = "Contested House Races")

table <- all_house_candiate_by_county %>%
  select(contest_name, choice_party, total_votes) %>%
  group_by(contest_name, choice_party) %>%
  summarize(Votes = sum(total_votes)) %>%
  ungroup() %>%
  mutate(
    contest_name = str_remove(contest_name, "US HOUSE OF REPRESENTATIVES "),
    choice_party = if_else(choice_party == "DEM" | choice_party == "REP", as.character(choice_party), "OTHER")
  )

ggplot(data = table, aes(x = contest_name, y = Votes, fill = choice_party)) +
  geom_col() +
  party_color_scale +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(x = "District", y = "Votes") +
  ggtitle("Results of All US House Races by Total Votes")

ggplot(data = table, aes(x = contest_name, weight = Votes, fill = choice_party)) +
  party_color_scale +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(x = "District", y = "Share of Votes") + geom_hline(yintercept = .5) +
  ggtitle("Results of All US House Races by Proportion of Vote Recieved")


ggplot(data = statewide_shares, aes(x = choice_party, y = party_total, fill = choice_party)) +
  geom_col() +
  party_color_scale + labs(x = "Party", y = "Votes") +
  ggtitle("Total Votes Recieved by Each Party Across All North Carolina US House Races")

ggplot(data = statewide_shares_no_uncontested) +
  geom_col(aes(x = choice_party, y = party_total, fill = choice_party)) +
  party_color_scale +
  labs(x = "Party", y = "Votes") +
  ggtitle("Total Votes Recieved by Each Party In Contested US House Races In North Carolina")
```

### Number of Mail Ballots Vs Number of Total Ballots

```{r MailRate}
ggplot(data = house_county_totals, aes(x = total_votes, y = mail)) +
  geom_point(
    data = house_county_totals %>%
      filter(
        county == "BLADEN",
        contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"
      ),
    aes(x = total_votes, y = mail), color = "yellow", size = 5
  ) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  scale_x_continuous(breaks = seq(0, 300000, by = 50000), labels = scales::comma) +
  xlab("Total Votes Cast") +
  ylab("Votes Cast by Mail") +
  ggtitle("Number of Mail in Ballots vs Total Number of Ballots")

ggplot(data = house_county_totals, aes(x = total_votes, y = mail)) +
  geom_point(
    data = house_county_totals %>%
      filter(
        county == "BLADEN",
        contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"
      ),
    aes(x = total_votes, y = mail), color = "yellow", size = 5
  ) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  scale_x_continuous(breaks = seq(0, 10000, by = 2000), labels = scales::comma) +
  xlab("Total Votes Cast") +
  ylab("Votes Cast by Mail") +
  coord_cartesian(xlim = c(0, 10000), ylim = c(0, 1000)) +
  ggtitle("Number of Mail in Ballots vs Total Number of Ballots in Small Counties")

ggplot(data = house_county_totals, aes(x = total_votes, y = mail / total_votes)) +
  geom_point(
    data = house_county_totals %>%
      filter(
        county == "BLADEN",
        contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"
      ),
    aes(x = total_votes, y = mail / total_votes), color = "yellow", size = 5
  ) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  scale_x_continuous(breaks = seq(0, 300000, by = 50000), labels = scales::comma) +
  scale_y_continuous(labels = scales::percent) +
  ylab("Percentage of Votes Cast by Mail") +
  xlab("Total Votes Cast in County") +
  ggtitle("Rate of Mail in Voting vs Total Number of Ballots")

ggplot(data = house_county_totals, aes(x = total_votes, y = mail / total_votes)) +
  geom_point(data = house_county_totals %>% filter(county == "BLADEN", contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"), aes(x = total_votes, y = mail / total_votes), color = "yellow", size = 5) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  scale_x_continuous(breaks = seq(0, 10000, by = 2000), labels = scales::comma) +
  scale_y_continuous(labels = scales::percent) + coord_cartesian(xlim = c(0, 10000)) +
  ylab("Percentage of Vote by Mail") +
  xlab("Total Votes Cast in County") +
  ggtitle("Rate of Mail in Voting vs Total Number of Ballots in Small Counties")
```

### Vote by Mail vs In Person Voting

```{r mail-in-vs-in-person, message = FALSE}
table3 <-
  house_by_county %>%
  mutate(in_person = total_votes - mail) %>%
  gather(
    key = vote_type,
    value = vote_count,
    total_votes,
    election_day,
    early,
    mail,
    provisional,
    in_person
  ) %>%
  filter(!choice == "county_total") %>%
  group_by(contest_name, county, vote_type) %>%
  mutate(county_total = sum(vote_count)) %>%
  mutate(share = vote_count / county_total) %>%
  select(-vote_count, -county_total) %>%
  filter(choice_party == "DEM" | choice_party == "REP") %>%
  spread(vote_type, share) %>%
  rename_at(vars(-contest_name, -county, -choice, -choice_party), function(x)(paste0(x, "_share")))

table3 %<>% full_join(house_county_totals) %>%
  filter(contest_name != "US HOUSE OF REPRESENTATIVES DISTRICT 03")

ggplot(data = table3, aes(x = in_person_share, y = mail_share, color = choice_party)) +
  geom_point(data = table3 %>% filter(county == "BLADEN", contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"), aes(x = in_person_share, y = mail_share), color = "yellow", size = 5) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) + geom_abline(intercept = 0, slope = 1) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_color_manual(values = setNames(c(blue, red), c("DEM", "REP"))) +
  xlab("In Person Votes") +
  ylab("Vote by Mail") +
  ggtitle("Partisan Gap Between Vote by Mail and In Person Voting")

ggplot(data = table3, aes(x = early_share, y = mail_share, color = choice_party)) +
  geom_point(data = table3 %>% filter(county == "BLADEN", contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"), aes(x = early_share, y = mail_share), color = "yellow", size = 5) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  geom_abline(intercept = 0, slope = 1) + scale_y_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_color_manual(values = setNames(c(blue, red), c("DEM", "REP"))) +
  xlab("Early Voting") +
  ylab("Vote by Mail") +
  ggtitle("Partisan Gap Between Vote by Mail and Early Voters")

ggplot(data = table3, aes(x = election_day_share, y = mail_share, color = choice_party)) +
  geom_point(data = table3 %>% filter(county == "BLADEN", contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"), aes(x = election_day_share, y = mail_share), color = "yellow", size = 5) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  geom_abline(intercept = 0, slope = 1) + scale_y_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_color_manual(values = setNames(c(blue, red), c("DEM", "REP"))) +
  xlab("Election Day") +
  ylab("Vote by Mail") +
ggtitle("Partisan Gap Between Vote by Mail and Election Day Voters")
```

### Early Voting vs Election Day Voting
```{r early-vs-election-day,results="hold"}
ggplot(data = table3, aes(x = election_day_share, y = early_share, color = choice_party)) +
  geom_point(data = table3 %>% filter(county == "BLADEN", contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"), aes(x = election_day_share, y = early_share), color = "yellow", size = 5) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  geom_abline(intercept = 0, slope = 1) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  scale_color_manual(values = setNames(c(blue, red), c("DEM", "REP"))) +
  ggtitle("Partisan Gap Between Early Voters and Election Day Voters")

DemLM <- lm(
  early_share ~ election_day_share,
  data = table3 %>% filter(choice_party == "DEM")
)
RepLM <- lm(
  early_share ~ election_day_share,
  data = table3 %>% filter(choice_party == "REP")
)
summary(DemLM) %>% coefficients() %>% kable(caption = "Democrat")
summary(RepLM) %>% coefficients() %>% kable(caption = "Republican")
```

### Partisan Gap Between Election Day and Early Voting

```{r early-vs-election-day-rate,results="hold"}
ggplot(data = table3, aes(x = election_day / total_votes, y = early_share - election_day_share, color = choice_party)) +
  geom_point(data = table3 %>% filter(county == "BLADEN", contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"), aes(x = election_day / total_votes, y = early_share - election_day_share), color = "yellow", size = 5) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) + geom_hline(yintercept = 0) +
  scale_y_continuous(labels = scales::percent_format()) + scale_x_continuous(labels = scales::percent_format()) +
  scale_color_manual(values = setNames(c(blue, red), c("DEM", "REP"))) +
  labs(x = "Percentage of Total Votes Cast on Election Day", y = "Share of Early Ballots vs Election Day Ballots") +
  ggtitle("Gap between Early and Election Day Voting vs Rate of Election Day Voting")

DemLM <- lm(
  I(early_share - election_day_share) ~ I(election_day / total_votes),
  data = table3 %>% filter(choice_party == "DEM")
)
RepLM <- lm(I(early_share - election_day_share) ~ I(election_day / total_votes),
  data = table3 %>% filter(choice_party == "REP")
)
summary(DemLM) %>% coefficients() %>% kable(caption = "Democrat")
summary(RepLM) %>% coefficients() %>% kable(caption = "Republican")
```

### Any Relationship Between Vote by Mail Prevalence and Partisan Lean

```{r mail-in-rate, results="hold",collapse=TRUE}
ggplot(data = table3, aes(x = mail / total_votes, y = mail_share, color = choice_party)) +
  geom_point(data = table3 %>% filter(county == "BLADEN", contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"), aes(x = mail / total_votes, y = mail_share), color = "yellow", size = 5) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format()) +
  geom_hline(yintercept = .5) +
  scale_x_continuous(labels = scales::percent_format()) +
  scale_color_manual(values = setNames(c(blue, red), c("DEM", "REP"))) +
  labs(x = "Percentage of Total Votes Cast by Mail", y = "Share of Mail In Ballots") +
  ggtitle("Percentage of Mail in Ballots Won vs Rate of Voting by Mail")

table3 <- table3 %>%
  mutate(mail_rate = mail / total_votes, mail_gap = mail_share - in_person_share)

graph <-
  ggplot(data = table3, aes(x = mail_rate, y = mail_gap, color = choice_party)) +
  geom_point(data = table3 %>% filter(county == "BLADEN", contest_name == "US HOUSE OF REPRESENTATIVES DISTRICT 09"), aes(x = mail_rate, y = mail_gap), color = "yellow", size = 5) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_continuous(labels = scales::percent_format()) +
  geom_hline(yintercept = 0) +
  scale_color_manual(values = setNames(c(blue, red), c("DEM", "REP"))) +
  xlab("Percentage of Total Votes Cast by Mail") +
  ylab("Difference Between Share of Mail In and In Person Ballots") +
  ggtitle("Rate of Voting by Mail vs Mail/In Person Partisan Gap")

DemLM <- lm(mail_gap ~ mail_rate, data = table3 %>% filter(choice_party == "DEM"))
RepLM <- lm(mail_gap ~ mail_rate, data = table3 %>% filter(choice_party == "REP"))
summary(DemLM) %>% coefficients() %>% kable(caption = "Democrat")
summary(RepLM) %>% coefficients() %>% kable(caption = "Republican")
graph
```
